How to write a PhD in Biological Sciences: a guide for the uninitiated

### 2023-04-29 1.01-beforeyoustart.Rmd
Add files via upload
# with addition of conceptual framework chapter

I prefer a thesis that fits together within an explicit [conceptual framework](#PhDintro). I like PhD students to conceptualise how their chapters are interlinked, and to present this as a figure in the introduction of their [proposal](#proposal). A useful concept to be aware of is the 'Hierarchy of Hypotheses' [@heger2018hierarchy]. In this approach, you are encouraged to consider the '[bigger question](#bigidea)' in your research area, identifying both what studies have produced compelling evidence, and identifying knowledge gaps from a theory driven approach. Generally, it is only possible to consider these types of concepts once you are familiar with the literature (both theoretical and experimental).

>>>>

I prefer a thesis that fits together within an explicit [conceptual framework](#concept). I like PhD students to conceptualise how their chapters are interlinked, and to present this as a figure in the [introduction](#PhDintro) of their [proposal](#proposal). A useful idea to be aware of is the 'Hierarchy of Hypotheses' [@heger2018hierarchy]. In this approach, you are encouraged to consider the '[bigger question](#bigidea)' in your research area, identifying both what studies have produced compelling evidence, and identifying knowledge gaps from a theory driven approach. Generally, it is only possible to consider these types of concepts once you are familiar with the literature (both theoretical and experimental).

### 2023-04-29 2.20-proposal.Rmd
Add files via upload
# with addition of conceptual framework chapter

@@

>>>>

A key part of the introduction will be your [conceptual framework](#concept) that will explain how your thesis chapters fit together, and crucially, add up to more than a sum of their parts.

@@

## Thesis timeline

You should be able to state from the outset when you are going to find the time to fit in all of the components of your thesis. Being able to plan when and how you will do the work is a key aspect of your proposal. I discuss how to construct a [timeline](#Gendt) in the form of a Gendt diagram in another chapter (see [Part II](#Gendt)).

### 2023-04-29 2.21-proposal-presentation.Rmd
Add files via upload
# with addition of conceptual framework chapter

1. You need a title slide (with your name, that of your advisor and institution)
2. Several slides of introduction
  + that put your study into the big picture
  + explain variables in the context of existing literature
  + explain the relevance of your study organisms
  + give the context of your own study
3. Your aims and hypotheses
4. Methods and Materials
  + Images of apparatus or diagrams of how apparatus are supposed to work. If you can’t find anything, draw it simply yourself.
  + Your methods can be abbreviated but the methods are important for a proposal, especially the numbers of replicates and the variables that you plan to measure from them. 
  + Analyses are important. Make sure that you understand how they work, otherwise, you won’t be able to present them to others. Importantly, explain where each of the variables that you introduced, and explained how to measure, fit into the analyses. There shouldn’t be anything new or unexpected that pops up here.
5. Expected results
6 like to see what the results might look like, even if you have to draw graphs with your own lines on it. Use arrows to show predictions under different assumptions.
7. Budget
8. Timeline

>>>>

1. You need a title slide (with your name, that of your advisor and institution)
2. Several slides of introduction
  + that put your study into the big picture
  + explain variables in the context of existing literature
  + explain the relevance of your study organisms
  + give the context of your own study
  + provide your [conceptual framework](#concept) for your thesis
3. Your aims and hypotheses
4. Methods and Materials
  + Images of apparatus or diagrams of how apparatus are supposed to work. If you can’t find anything, draw it simply yourself.
  + Your methods can be abbreviated but the methods are important for a proposal, especially the numbers of replicates and the variables that you plan to measure from them. 
  + Analyses are important. Make sure that you understand how they work, otherwise, you won’t be able to present them to others. Importantly, explain where each of the variables that you introduced, and explained how to measure, fit into the analyses. There shouldn’t be anything new or unexpected that pops up here.
5. Expected results
6 like to see what the results might look like, even if you have to draw graphs with your own lines on it. Use arrows to show predictions under different assumptions.
7. Budget
8. Show your [timeline](#Gendt)
